Challenges to transboundary fisheries management in North America under climate change

### 2020-11-21 Ecol/Supplements.Rmd
Update to manuscript

---
output:
  bookdown::pdf_book:
    toc : false
  # pdf_document: default
  # word_document: default
editor_options:
  chunk_output_type: console
csl: ../References/E&S.csl
bibliography: ../References/Oceans_Canada_Biblio.bib
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
---

```{r setup_oc, eval = T, echo = F, warning = F, message = F, results = 'hide'}

#### READ ME !!! ####
# Run this chunk before knit so you make sure you have all pkgs installed in R

library(MyFunctions)

#### Library ####
packages <- c(
 "readxl", # Read dataframe
 "dplyr", # Data manipulation
 "tidyr", # Data manipulation
 "ggplot2", #Nice grpahs and spatial analysis
 "cowplot",
 "rgdal",
 "RColorBrewer",
 "knitr",
 "kableExtra",
 "data.table",
 "ggrepel",
 "gridExtra",
 "ggmap",
 "rgeos",
 "sf",
 "sp",
 "rgdal", #Spatial analysis 
 "tools", #Spatial analysis 
 "png", # For reading plots in chunk codes
 "grid", # For reading plots in chunk codes
 "here"
)

my_lib(packages)

##________________________________________________##

 # Current time
Beg_In <- paste(2005); Beg_In
Beg_End <- paste(2014); Beg_End

# Mid century
Mid_In <- paste(2041);Mid_In # 2046
Mid_End <- paste(2060);Mid_End

# End Century
End_In <- paste(2080);End_In # 2090
End_End <- paste(2099);End_End
```

<!-- ## Abstract -->

<!-- Climate change is shifting the distribution of fish stocks that straddle between Exclusive Economic Zones (EEZ), challenging transboundary fisheries management. Here, we examine the projected sharing of jointly managed transboundary fish stocks between Canada and the United States. We hypothesize that ocean warming will alter the sharing of fish stocks between the two countries, and that such changes will intensify under a high carbon emission scenario. We look at the specific cases of the International Pacific Halibut Commission that manages Pacific halibut (*Hippoglossus stenolepis*) and a resource sharing arrangement in the Gulf of Maine for cod (*Gadus morhua*), haddock (*Melanogrammus aeglefinus*) and yellowtail flounder (*Limanda ferruginea*) to discuss the management consequences of shifts in transboundary stocks. We rely on multiple Earth system models’ simulations and species distribution models to estimate the change in catch potential and stock share ratio of each transboundary stock in the $21^{st}$ century under two climate change scenarios. Results show that, even under a low emission scenario, most transboundary fish stocks sharing ratios, i.e., the proportion of the total catch of a fish stock taken by a given country, will change by 2050 relative to present. The overall reduction in catch potential, in addition to the changes in stock-share will further exacerbate trade-offs between changes in species catch potential. Such trade-offs in the Atlantic and Pacific regions will be amplified if a high emission scenario is followed, relative to a low carbon emission scenario. Based on the simulation results, we examine possible solution options to reduce climate risks on transboundary fish stocks and fisheries. -->

<!-- Keywords: *Transboundary fisheries management; climate change; joint fisheries management; species distribution shift* -->

## Introduction

<!-- ##### First paragraph of paper summary -->

In 1982, the United Nations Law of the Sea Convention (UNCLOS) formalized the concept of exclusive economic zones (EEZs) creating what we know today as shared stocks [@UnitedNations:1986tl], i.e. stocks that migrate between countries EEZs (known as transboundary stocks) or between EEZs and the high seas (also called straddling stocks) [@Song:2017iua]. Today, an estimated 347 [@Teh:2015gd] to 1500 [@Caddy:1997ue] fish stocks cross national borders, some of them jointly managed by two or more countries. These stocks are responsible for almost 50% of these countries total fish catches [@Teh:2015gd]. Under Article 63, UNCLOS incentives actions to cooperate on the management of shared stocks [@UnitedNations:1986tl] as often management success depends on effective cooperation between parties [@Munro:2002uf; @Sumaila:2013hv]. Since the definition of shared stocks, game theory has been one of the most common approaches used to analyze the management of these type of stocks. However, shared stocks' management can be convoluted due to the participation of several fishing “players”, different countries and sometimes jurisdictions within a country, the migration patterns of the stock, and their abundance fluctuation within space and time [@Munro:2002uf, Engler and Saunders *in press*]. In addition, international treaties might not be prepared to address the effects that climate change will bring to shared fish stocks (Engler and Saunders *in press*, Koubrak and VanderZwaag, *in press*).

<!-- ##### Climate Change and Transboundary Fisheries -->

The ocean is getting warmer [@Rheim:2013fv], less oxygenated [@Schmidtko:2017cl], and increasing in acidity [@Ross:2011kd; @IPCC:2019tn]. To cope with these changes in ocean biophysic properties, marine species, including shared fish stocks, have been shifting their distribution towards the poles and/or deeper waters [@Poloczanska:2016kk]. As climate change reshapes the ocean's environment worldwide [@Gattuso:2015jz], shared fisheries' delicate governance is threatened as new migration patterns may arise [@Miller:2013iv; @Pinsky:2018cb], historic distribution and abundances might shift [@Cheung:2010dt], and species basic natural traits may modify [@Pauly:2017hpa]. Catches of shared stocks like tunas, have significantly increased in some regions such as the subtropical Atlantic and western Pacific Oceans and are projected to continue [@MonllorHurtado:2017cm; @ErauskinExtramiana:2019fw]. Multiple shared species in North America have been observed to shift in distribution following changes in optimal conditions such as sockeye salmon (*Oncorhynchus nerka*) [@McDaniels:2010vn], Atlantic cod [@Pershing:2015gq], and flounders [@Pinsky:2012kq]. Moreover, these shifts are projected to continue towards the end of the $21^{st}$ century [@Cheung:2018hz]. As a result, some countries or management jurisdictions may see more shared fisheries and their catches shifting into their waters while others will stand to lose [@Oremus:2020id; @Pinsky:2018cb]. Nevertheless, management rules for shared stocks (e.g. quota or spatial delimitation) are often determined based on current and/or historic knowledge of the stock's distribution and do not consider future shifts in distributions [@FredstonHermann:2018kp].

<!-- ##### Past Conflicts of shifting stocks -->

The shifts in distribution of shared fish stocks will impact the economics of their fisheries [@Pinsky:2012kq; @Sumaila:2019gh; @Sumaila:2020tp], and create international disputes between countries [@Munro:2002uf; @Spijkers:2017ij; @Pinsky:2018cb]. Canada and the US share important transboundary stocks of salmon (*Oncorhynchus spp.*), Pacific halibut (*Hippoglossus stenolepis*), and Atlantic cod (*Gadus morhua*) offering a unique lens to understand the extent to which climate-induced distributional shifts will challenge the future sustainability of transboundary fisheries. These countries have a long history of fisheries cooperation participating in diverse, jointly managed, commercial transboundary stocks through various fisheries management organizations [@NOAA:2018tu]. Furthermore, climate-related fluctuations in stocks' distribution have historically created disputes between Canada and the US, increasing international conflict and threatening the health of diverse transboundary stocks [@Munro:2002uf; @CIA:2017uta].

It is expected that climate induced shifts in stocks’ distribution will affect the rules in place that keep international treaties alive. Therefore, the main objective of this article is to assess the level of exposure that bi-lateral transboundary fisheries treaties between Canada and the US have to climate change through shifts in stock distributions. Specifically, we rely on a species distribution model and scenario planning to project the changes in the distribution of selected fish stocks jointly managed by Canada and the United States focusing on two case studies (the International Pacific Halibut Commission and a fisheries arrangement for the Gulf of Maine) Finally, we explore similar situations around the world and identify opportunities to improve the adaptability of transboundary stocks management to climate change in North America. Despite an overall expectation of species following a poleward shift, important geographic constraints (e.g. Gulf of Alaska representing a latitudinal block) [@Kleisner:2016je], geo-political features (e.g. the localization of Alaska in reference to Canada and the contiguous states), and management rules (e.g. quota allocations, spatial management rules) may play an important role in the redistribution of benefits. Understanding these stocks shifts will shed a light on future conditions and inform decision-makers on the paths to follow under a changing climate.

## Materials and Methods

### Study area and fisheries

The current study used the International Pacific Halibut Commission (IPHC) and the Gulf of Maine arrangement (hereafter referred as GoMA) as case studies to discuss the implications that climate change could have in the management of transboundary stocks. For the IPHC, we used the most updated spatial regulatory data along its 12 regulatory areas [@IPHCSecreatriat:2017uz; @IPHC:2019wt]. For this specific case, we considered Alaska as a separate entity, the US contiguous states as a second one (Washington, Oregon and California), and lastly British Columbia (Canada). For the GoMA we used the Northwest Atlantic Fisheries Organization's (NAFO) divisions 5Y, 5Ze, and 4X[^NAFO] within latitudes 46.2°N and 41.5°S, and longitudes -72°W and -64°E (Figure \ref(fig:ocFigure1-map)). Is worth mention that, while NAFO's divisional zones were used in this study for domestic management, NAFO does not manage fisheries within the EEZs of Canada and the United States. Fisheries data fpr the projections was gathered from the Sea Around Us from 1951 to 2014 [@Zeller2016].

[^NAFO]: Northwest Atlantic Fisheries Organization, Available at https://www.nafo.int/Science/

```{r Study_Area_Map, eval = F, echo = F, warning=F, message=F}

Countries <- c("Canada","United States")

### North America Initial Map ####
path_world <- my_path("G",Extra_Path = "Spatial/TM_WORLD_BORDERS-0.3")
file_name <- "TM_WORLD_BORDERS"

World_Land_sf <- st_read(dsn = path_world, 
layer = file_path_sans_ext(file_name)
) %>% 
st_transform(crs = 3832)

# names(World_Land_sf)

North_America_Land <- World_Land_sf %>% 
filter(NAME %in% Countries) %>% 
st_transform(crs = 3832)

# head(North_America)

# ggplot(North_America_Land) +
#  geom_sf() +
#  coord_sf(ylim = c(3051701,11528813)
#  )

### North America EEZs Map ####
path_eez <- my_path("G",Extra_Path = "Spatial/World_EEZ_v10_2018-0.3")
#The File
fnam_eez <- "eez_v10.shp"

eez_world_sf <- st_read(dsn = path_eez, 
            layer = file_path_sans_ext(fnam_eez)
) %>% 
 st_transform(crs = 3832)

Nort_America_EEZ <- eez_world_sf %>% 
filter(Territory1 %in% Countries)

# Testing
ggplot(Nort_America_EEZ) +
geom_sf()

# Testing together

ggplot(North_America_Land) +
 geom_sf() +
 geom_sf(data = Nort_America_EEZ, colour = "black") +
 coord_sf(ylim = c(3051701,11528813)
 )

#### IPHC Regulatory Areas ####

path_IPHC <- my_path("D",Extra_Path = "IPHC/IPHC_RegulatoryAreas/")

# The Canada file
fnam_IPHC_C <- "IPHC_RegulatoryAreas_CAN"

IPHC_Reg_C <- st_read(dsn = path_IPHC, 
           layer = file_path_sans_ext(fnam_IPHC_C)
) %>% 
 select(
  ID,
  Country = COUNTRY,
  State = PROV_EN,
  RegArea
 ) %>% 
 mutate(Comments = "NA") %>% 
 st_transform(crs = 3832)

# 4326 use this to merge with DBEM data
# 3832 use this to center the map and plot

# ggplot(IPHC_Reg_C) +
# geom_sf()

# The US File
fnam_IPHC_US <- "IPHC_RegulatoryAreas_US"

IPHC_Reg_US <- st_read(dsn = path_IPHC,
            layer = file_path_sans_ext(fnam_IPHC_US)
) %>%
 st_transform(crs = 3832)

# head(IPHC_Reg_US)
ggplot(IPHC_Reg_US) +
 geom_sf()

### Combine both shapefiles

IPHC_Reg <- rbind(IPHC_Reg_US,IPHC_Reg_C)

ggplot(IPHC_Reg) +
 geom_sf()

##### NAFO Regulatory Area ####

path_NAFO <- paste(Data_path,"Spatial/NAFO")
#The File
fnam_NAFO <- "Divisions"

NAFO_Area <- st_read(dsn = path_NAFO,
           layer =file_path_sans_ext(fnam_NAFO))%>% 
 st_transform(crs = 3832)

Zones <- c(
 "5Ze",
 # "5Zw",
 "5Y",
 "4X"
)

NAFO_5z <- NAFO_Area %>% 
filter(ZONE %in% Zones)

# head(NAFO_5z)

ggplot() +
geom_sf(data = North_America_Land,
    fill = "lightgrey",
    colour = "black") +
 my_ggtheme_m()()

#### The final Map ####

# Homonize collors 
brewer.pal(name="Set2",5)

# Canada "#grey80"
# US "#grey40", 
# NAFO "#66CC99"
# IPHC #f01282

#North America point of view

North_A <- ggplot() +
 geom_sf(data = North_America_Land, 
     aes(fill = NAME),
     colour = "black"
 ) +
 geom_sf(data = IPHC_Reg, 
     aes(fill = "IPHC") ,
     colour = "black"
 ) +
 geom_sf(data = NAFO_5z, 
     aes(fill = "NAFO"),
     colour = "black"
 ) +
 my_ggtheme_m()() +
 scale_fill_manual("Region",
          breaks = c("Canada","IPHC","NAFO", "United States"), 
          values=c("grey80","#CC6666","#66CC99","grey40")
          ) +
 theme(legend.position = "right")

ggsave("Fig1_NA.png",
    plot = North_A,
    width = 6,
    height = 5,
    units = "in",
    path = Results_Path)

#NAFO #

NAFO <- ggplot() +
 geom_sf(data = NAFO_5z, #NAFO regions
     aes(fill = "NAFO"),
     colour = "black"
 )+
 geom_sf_label(data = NAFO_5z, # RegAreas labels
        aes(
         label = ZONE#,
         # fill = "NAFO"
        ),
        alpha = 0.5
 ) +
 geom_sf(data = North_America_Land, # North America Land
     aes(fill=NAME),
     colour = "black"
 ) +
 coord_sf(ylim = c(5951701,4728813),
      xlim = c(15373019,16573019)
 ) +
 # scale_fill_brewer("Region",palette = "Set2") +
 theme(
  panel.background = element_blank(),
  panel.border   = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_line(color = "transparent"),
  legend.position = "bottom",
  axis.title = element_blank()
 ) +
 scale_fill_manual("Region",breaks = c("Canada", "NAFO", "United States"), 
          # values = c("#E78AC3","#A6D854","8DA0CB")
          values=c("grey80", "#66CC99", "grey40")
          ); NAFO

# Save plot for manuscrip
ggsave("Fig1_NAFO.png",
    plot = NAFO,
    width = 6,
    height = 5,
    units = "in",
    path = my_path("R"))

# IPHC

IPHC_Reg$RegArea <- gsub("Closed Area","CA",IPHC_Reg$RegArea)

IPHC <- ggplot() +
 geom_sf(data = IPHC_Reg,
     aes(fill = "IPHC"),
     colour = "black"
 ) + 
 geom_sf(data = North_America_Land, # North America Land
     aes(fill=NAME),
     colour = "black"
 ) +
 geom_sf_label(data = IPHC_Reg, # RegAreas labels
        aes(
         label = RegArea#,
         # fill = Percentage
        ),
        alpha = 0.5
 ) +

 coord_sf(xlim = c(1963800, 10041910), # Subsets north america
      ylim = c(3651701,9728813)
 ) +
 theme(
  panel.background = element_blank(),
  panel.border   = element_blank(),
  panel.grid.minor = element_blank(),
  panel.grid.major = element_line(color = "transparent"),
  legend.position = "bottom",
  axis.title = element_blank()
 ) +
 scale_fill_manual("Region",breaks = c("Canada", "IPHC", "United States"), 
          values=c("grey80", "#CC6666", "grey40")
          )

# Save plot for manuscrip
ggsave("Fig1_IPHC.png",
    plot = IPHC,
    width = 6,
    height = 5,
    units = "in",
    path = my_path("R"))

```{r ocFigure1-map, eval = T, echo=FALSE, fig.height = 4, fig.width = 6, fig.cap="Map of Canada and the US with the regulatory areas of the International Pacific Halibut Commission and the NAFO sub-divisions containing the Gulf of Maine arrangement"}

img <- readPNG("./Figures/Figure 1.png")
grid.raster(img)

```{r Spatial_data, eval = F, echo = F, warning=F, message=F, results = 'hide'}

# List of DBEM INDEX within the EEZ
#Data provided by Vicki Lam
EEZIDs_List <- my_path("G",Extra_Path = "Spatial/DBEM",Name="Updated_EEZList_17June2016.xlsx",Read = TRUE)

EEZ_CellID <- my_path("G",Extra_Path = "Spatial/DBEM",Name="EEZ_CellID.xlsx",Read = TRUE)
colnames(EEZ_CellID) <- c("EEZID","INDEX")

# DBEM Coordinate system
Coor <- my_path("G",Extra_Path = "Spatial/DBEM", Name = "Lon_Lat_DBEM.txt", Read = TRUE)
colnames(Coor) <- c("INDEX","Longitude","Latitude")

### Select EEZs
EEZs <- c(
 "Canada (Pacific)",
 "Canada (Arctic)",
 "USA (West Coast)",
 "USA (Alaska, Arctic)",
 "USA (Alaska, Subarctic)",
 "Canada (Arctic)",
 "Canada (East Coast)",
 "USA (East Coast)"
)

### North America Initial Map ####

path_world <- my_path("D",Extra_Path = "Spatial/TM_WORLD_BORDERS")
file_name <- "TM_WORLD_BORDERS"

World_Land_sf <- st_read(dsn = path_world, 
             layer = file_path_sans_ext(file_name)
) %>% 
 st_transform(crs = 4326)

# names(World_Land_sf)

North_America_Land <- World_Land_sf %>% 
 filter(NAME %in% c("Canada","United States")) %>% 
 st_transform(crs = 3832) # 4326 use this to merge with DBEM data
# 3832 use this to center the map and plot

#### _____________________TEST _____________________ #
# head(North_America_Land)
# ggplot(North_America_Land) +
#  geom_sf() +
#  coord_sf(xlim = c(-120, -179),
#      ylim = c(30,70)
#  )
##__________________________________________ #

### North America EEZs Map ####

path_eez <- my_path("D",Extra_Path = "Spatial/World_EEZ_v10_2018")
#The File
fnam_eez <- "eez_v10.shp"

eez_world_sf <- st_read(dsn = path_eez, 
            layer = file_path_sans_ext(fnam_eez)
) %>% 
 st_transform(crs = 4326)# 4326 use this to merge with DBEM data
# 3832 use this to center the map and plot

#### _____________________TEST _____________________ #
# head(eez_world_sf)
### Test this map
# ggplot(ezz_North_America) +
#  geom_sf()
##__________________________________________##

#### IPHC Regulatory areas ####

## Correct DBEM for Alaska ####
IHPC_EEZs <- c(
 "Canada (Pacific)",
 # "Canada (Arctic)"
 "USA (West Coast)",
 "USA (Alaska, Arctic)",
 "USA (Alaska, Subarctic)"#,
 # "Russia (Far East)"
)

IPHC_EEZsID <- EEZIDs_List %>% 
 filter(Name %in%IHPC_EEZs) %>% 
 left_join(EEZ_CellID,
      by = "EEZID") %>% 
 left_join(Coor,
      by = "INDEX")

# "Move" the map so results are not in the other side of the world and it matches the projection I want
IPHC_Coor_Corrected <- IPHC_EEZsID %>% 
 filter(Longitude > 0) %>% 
 mutate(Longitude = Longitude-360)

IPHC_Coor_Corrected <- IPHC_EEZsID %>% 
 filter(Longitude < 0) %>% 
 bind_rows(IPHC_Coor_Corrected) %>% 
 filter(Latitude > 25)

# Create new column to merge with shapefiles

IPHC_Coor_Corrected_Clean <- IPHC_Coor_Corrected %>% 
 mutate(
  State = ifelse(Name == "Canada (Pacific)", "British Columbia",
          ifelse(Name == "USA (West Coast)","WAORCA","Alaska"))
 ) %>% 
 select(
  # State,
  INDEX,
  longitude = Longitude,
  latitude = Latitude)

# Now we load IPHC shapefiles ####
path_IPHC <- my_path("D",Extra_Path = "Spatial/IPHC_RegulatoryAreas")
# The Canada file
fnam_IPHC_C <- "IPHC_RegulatoryAreas_CAN"

IPHC_Reg_C <- st_read(dsn = path_IPHC, 
           layer = file_path_sans_ext(fnam_IPHC_C)
) %>% 
 select(
  ID,
  Country = COUNTRY,
  State = PROV_EN,
  RegArea
 ) %>% 
 mutate(Comments = "NA") %>% 
 st_transform(crs = 4326)

# 4326 use this to merge with DBEM data
# 3832 use this to center the map and plot

#### _____________________TEST _____________________##
# head(IPHC_Reg_C)
# ggplot(IPHC_Reg_C) +
# geom_sf()
##_________________________________________________##
# The US File
fnam_IPHC_US <- "IPHC_RegulatoryAreas_US"

IPHC_Reg_US <- st_read(dsn = path_IPHC, 
            layer = file_path_sans_ext(fnam_IPHC_US)
) %>% 
 st_transform(crs = 4326)

#### _____________________TEST _____________________ #
# head(IPHC_Reg_US)
# ggplot(IPHC_Reg_US) +
#  geom_sf()
##_________________________________________________##

### Combine both shapefiles

IPHC_Reg <- rbind(IPHC_Reg_US,IPHC_Reg_C)

#### _____________________TEST _____________________ #
# head(IPHC_Reg,12)
# tail(IPHC_Reg)
# unique(IPHC_Reg$State)

# It's going to take a while... 
# ggplot(IPHC_Reg) +
#  geom_sf()
##_________________________________________________##

# Now join the Shapefile with the DBEM lat/long dataset

# Convert DBEM to sf
IPHC_Coor_sf = st_as_sf(IPHC_Coor_Corrected_Clean,
            coords = c("longitude", "latitude")) %>% 
 st_set_crs(4326) %>% # to match shapefiles
 st_transform(4326) # to match shapefiles

# head(IPHC_Coor_sf)

# Merge DBEM sf to polygons
IPHC_Reg_DBEM <- st_join(IPHC_Reg,
             IPHC_Coor_sf, join = st_contains) %>% 
 filter(!is.na(INDEX)) %>%
 arrange(INDEX) 
# %>% 
#  st_transform(3832) # Go back to original projection otherwise doesnt like to plot

IPHC_Reg_DBEM_df <- as.data.frame(IPHC_Reg_DBEM) %>% 
 select(INDEX,RegArea)

#### _____________________TEST _____________________ #
# head(IPHC_Reg_DBEM_df)
# IPHC_Reg_DBEM_df %>% 
#  arrange(INDEX) %>% 
# head()
##_________________________________________________##

### STOP 
## Go to the results and now estimate the %change by region using IPHC_Reg_DBEM_df then, convert it to sf and merge it with IPHC_Reg

# Gulf of Maine ####
#### NAFO data ####

# path_NAFO <- paste(Data_Path,"/Spatial/NAFO",sep="")
#The File
# fnam_NAFO <- "Divisions"

# NAFO_Area <- st_read(dsn = path_NAFO,
# layer =file_path_sans_ext(fnam_NAFO))

Zones <- c(
 "5Ze",
 # "5Zw",
 "5Y",
 "4X"
)

# NAFO_5z <- NAFO_Area %>% 
# filter(ZONE %in% Zones)

#### _____________________TEST _____________________ #
# head(NAFO_5z)

# ggplot(NAFO_5z) +
# geom_sf()
# geom_sf(aes (fill = ZONE))
##_________________________________________________##

### Atlantic EEZ polygon (I don't think we need this for now) ####

Atlantic_EEZs <- c(
 "Canada (Arctic)",
 "Canada (East Coast)",
 "USA (East Coast)"
)

Atlantic_EEZsID <- EEZIDs_List %>%
 filter(Name %in% Atlantic_EEZs) %>%
 left_join(EEZ_CellID,
      by = "EEZID") %>%
 left_join(Coor,
      by = "INDEX")

### Atlantic DBEM Data ####

GB_Coor <- Coor %>% 
 filter(INDEX %in% Atlantic_EEZsID$INDEX) %>% 
 mutate(one = 1)

# Atlantic Land polygon ####

Countries <- c(
 "Canada",
 "United States"
)

### EEZ GM ####
eez_GM <- eez_world_sf %>% 
 filter(Territory1 %in% Countries)

#### _____________________TEST _____________________ #
# ggplot() +
#  geom_sf(data = eez_GM) +
#  coord_sf(xlim = c(-71.0592, -63.33333),
#      ylim = c(40.5,45.9759)
#      )
##_________________________________________________##

GM_Land <- World_Land_sf %>% 
 filter(NAME %in% Countries)

### The International Pacific Halibut Commission

The IPHC was established by Canada and the United States to oversee the management of Pacific halibut (*Hippoglossus stenolepis*) [@IPHC:2014tz]. There are 12 regulatory areas from which 3AB holds 51.2% of the stock, followed by regions 2ABC and 4ACDE with 23.1% and 20.4%, respectively, and lastly region 4B with only 5.2% of the stock distribution [@IPHC:2018vz]. In terms of management, the IPHC implements a total allowable catch (TAC) based on a yearly sampling of the Convention area in addition to a series of regulations to control fishing effort [@IPHC:2018vz]. The TAC is divided between recreational, subsistence and commercial fishery, with a portion set aside for bycatch of other fisheries [@IPHC:2019wt]. The commercial fishing season starts in March ending around November with restrictions allowing only set line gear with J-type hooks targeting individuals over 81.3 cm of total length [@IPHC:2019wt].

### The Gulf of Maine Arrangement

Since 1998 Canada and the US have used a "Resource Sharing Understanding" to inform the management of Eastern George Bank's Atlantic cod (*Gadus morhua*) and haddock (*Melanogrammus aeglefinus*) and, yellowtail flounder (*Limanda ferruginea*) [@Pudden:2007bo; @TRAC:2016wu; @Song:2017va]. From 2010 onward, the GoMA suggests catch-limits based on a weighted method where 10% represents the stocks’ historical distribution (from 1967 to 1994) and 90% current distributions produced by quarterly surveys and catch [@TRAC:2016wu]. Since its introduction, the average quota allocation for each species proposed by the GoMA has been (Table \@ref(tab:ocTable1); Atlantic cod 77% Canada and 23% US, haddock 55% Canada and 45% US, and yellowtail flounder 34% Canada and 66% US [@TRAC:2015wj; @TRAC:2015vp; @TRAC:2015tt; @TRAC:2018tw; @TRAC:2018tj; @TRAC:2018wx; @Trinko:2019ou]. However, because this is an unofficial agreement, Canada and the US ultimately take single management decisions [@Soboil:2006gw]. In terms of management, the US has a multi species harvest control with area and season closures, mesh sizes, effort control, and mobile gear vessels that use bottom otter trawl gear [@Soboil:2006gw]. In contrast, Canada has a quota system in addition to limited-entry licensing, fleet allocations, and mesh and fish size regulation, among other input controls. Canada inshore vessels fish cod with longline and gillnet while haddock is mainly caught with bottom otter trawl gear [@Soboil:2006gw].

```{r ocTable1, eval = T, echo = F, warning=F, message=F}

table_1 <- read_excel("./Tables/Tables.xlsx") 
colnames(table_1) <- c("Year","Canada","US","Canada","US","Canada","US")

table_1 %>% 
  kable("latex",
        booktabs = T,
        caption = "Historic quota allocation suggested for Atlantic cod, haddock and yellowtail flounder in the Gulf of Maine "
  ) %>% 
  kable_styling(full_width = F,
                position = "center") %>%
  add_header_above(c(" ", "Atlantic cod" = 2, "Haddock" = 2, "Yellowtail flounder" = 2)
  ) %>% 
  footnote(general = c("* Average from 2006 to 2009, before the implementation of the weighted method",
                       "' Since the implementation of the weighted method",
                       "'' Standard deviation of the wiighted average",
                       "^ TRAC (2015abc)",
                       "† TRAC (2018abc)",
                       "‡ Lake (2019)"
  )
  )

### Projecting future species distributions

We used a Dynamic Bioclimatic Envelope Model (DBEM) to project the distribution of species from 2015 to 2100 under two scenarios of climate change [@Cheung:2010dt; @Cheung:2016jd]. The DBEM algorithm integrated ecophysiology and habitat suitability with spatial population dynamics of exploited fishes and invertebrates to project shifts in abundance and potential fisheries catches under climate change. The algorithm predicted species distribution based on depth and latitudinal range, habitat preferences and an index of species association with major habitat types to estimate changes in abundance distribution over a $0.5^{\circ}$ x $0.5^{\circ}$ grid of the world ocean. For each grid cell and time step, the model then calculated species carrying capacity according to sea surface temperature, salinity, oxygen content, sea ice extent (for polar species) and bathymetry, as well as the species preferences to these conditions. It then incorporated the intrinsic population growth, settled larvae, and net migration of adults from surrounding cells using an advection-diffusion-reaction equation. Finally, the model also simulated the effects of changes in temperature and oxygen content on growth of individuals [@Cheung:2013gk; @Cheung:2016jd]. Ultimately, the model simulated spatial and temporal population dynamics, and estimated a proxy of maximum sustainable yield (MSY) by applying fishing at MSY level for each grid cell, hereafter referred as maximum catch potential (MCP).

We projected the DBEM using three Earth system models (ESM), the Geophysical Fluid Dynamics Laboratory Earth System Models 2M (GFDL)[^GFDL], the Institute Pierre Simon Laplace Climate Model 5 (IPSL-CM5)[^IPSL], and the Max Planck Institute for Meteorology Earth System Model (MPI)[^MPI]. Each model was downscaled to match the DBEM $0.5^{\circ}$ x $0.5^{\circ}$ grid using the nearest neighbor method, and in some cases, bilinear interpolation [@Cheung:2016hf]. Finally, we used the model outputs for two scenarios of the Intergovernmental Panel on Climate Change (IPCC)-Representative Concentration Pathways (RCP) 2.6 and 8.5 representing a low greenhouse gas emission (strong mitigation) and a high greenhouse gas emission (week mitigation) scenario, respectively [@IPCC:2014bz]. To estimate model robustness and capture the structural uncertainty build within ESM models, we averaged the DBEM results for all three models ($\mu \pm \sigma$) and marked regions where at least one ESM disagree in direction with the rest.

[^GFDL]: Geophysical Fluid Dynamics Laboratory Earth System Models 2M [online] https://www.gfdl.noaa.gov
[^IPSL]:Institute Pierre Simon Laplace Climate Model 5 [online] https://www.icmc.ipsl.fr
[^MPI]: Max Planck Institute for Meteorology Earth System Model [online] https://www.mpimet.mpg.de/en/science/models

```{r Data_Loading_For_DBEM, eval = F, echo = F, warning = F ,message = F}

# Species data (Names, treaties, etc)
# DBEM exploited species list (ONLY ON DROBO)
exploited_species_list <- fread("/Volumes/DATA/PROJECTION exploited species ENM/exploited_species_list.csv")

# IPHC and GB species
Selected_Species <- fread(paste(Data_Path,"Species/List_Species.csv",sep=""))

# NAFO Species
# This is the subset of species Fished on canada and US waters of all NAFO's species list
NAFO_Species <- read.csv(paste(Data_Path,"Species/NAFO_Species.csv",sep="")) %>%
 mutate(Treaty = "NAFO")

```{r DBEM_Data_For_All, eval = F, echo = F, warning = F, message = F}

#-------------------- NOTE --------------------#
# Chunk to be ran in Hall100 connected to DROBO
#---------------------------------------------#

All_Species <- exploited_species_list %>% 
 # filter(TaxonName %in% c("Hippoglossus stenolepis","Anoplopoma fimbria")) %>% 
 filter(TaxonName %in% NAFO_Species$TaxonName) %>%
 mutate(Treaty = "NAFO") %>% 
 bind_rows(Selected_Species) %>% 
 filter(! duplicated(TaxonName)) %>% # Removes duplicated
 select(TaxonName,
     TaxonKey,
     CommonName,
     Treaty)

# For suplement Material ####
# Suplemental_table <- All_Species %>% 
#  select(-TaxonKey) %>% 
#  mutate(Treaty = ifelse(TaxonName %in% NAFO_Species$TaxonName,"NAFO",
#             ifelse(TaxonName %in% c("Hippoglossus stenolepis",
#                         "Anoplopoma fimbria"),
#                 "IPHC",
#                 "PSC")
#             )
#  ) %>% 
#  rename("Scientific name" = TaxonName,
#     "Common name" = CommonName) 
# 
#  write.csv(Suplemental_table,
#       "Final_Species_List.csv",
#       row.names = F)

#_____________________________END SUPLEMENTAL TABLE_______________________________#

Models <- c(
 "GFDL26F1",
 "GFDL85F1",
 "IPSL26F1",
 "IPSL85F1",
 "MPI26F1",
 "MPI85F1"
)

for(m in 1:length(Models)){
 # for(s in 1:2){ #For tensting models
 for(s in 1:nrow(All_Species)){

  # Imports all data once#

  x <-dbem_Import(All_Species[s,2],# Include Treaties information
          2005,
          2099,
          Model = Models[m],
          Data_Type= "Catch"
  )

  # } # For testing the dbem_Import function

  ### If statement for abstent species ##

  if(length(x >= 1)){

   # Merging with lat/long ###
   Spp_Pro <- x %>% 
    left_join(Coor,
         by = "INDEX")

   # If statement for those species not found within national EEZ
   if(nrow(Spp_Pro) <= 1){

    print("No data within Area Selected")

   }else{

    # Include Species Name
    Spp_Pro$Species <- paste(All_Species[s,1])
    # Include Treaties information
    Spp_Pro$Treaty <- paste(All_Species[s,4])

    Spp_Pro_I <- as.data.table(Spp_Pro)

    # Groups all species together in one database
    if(s == 1){
     Data_P <- copy(Spp_Pro_I) # <- copies the previouse data.table
    }else{
     setkey(Data_P, # <- sets the data.table as "reference" ?setkey
         Species); setkey(Spp_Pro_I,
                 Species);
     list <- list(Data_P, # <- creates a list to merge the tables
            Spp_Pro_I)

     Data_P <- rbindlist(list, #<- Merges all the data in one single file.
               use.names = TRUE, # <- This will merge columns by Name
               fill = TRUE, # <- This allows for the NA's
               idcol = NULL) # Columns id
    }
   }
  } #if statement for species outside EEZ
 } # Species loop end

 #Include wich model the data is from
 Data_P$Model <- paste(Models[m])

 # Groups all models together in one database
 if(m == 1){
  DBEM_Results <- copy(Data_P) # <- copies the previouse data.table
 }else{
  setkey(DBEM_Results, # <- sets the data.table as "reference" ?setkey
      Species); setkey(Data_P,
              Species);
  list <- list(DBEM_Results, # <- creates a list to merge the tables
         Data_P)

  DBEM_Results <- rbindlist(list, #<- Merges all the data in one single file.
               use.names = TRUE, # <- This will merge columns by Name
               fill = TRUE, # <- This allows for the NA's
               idcol = NULL) # Columns id
 }

} #End of loops

>>>>

---
output:
  bookdown::pdf_book:
    toc : no
    keep_tex: no
    number_sections: yes
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
- \floatplacement{table}{H}
- \usepackage{caption}
- \captionsetup[figure]{name = Figure A4}
- \captionsetup[table]{name = Table A4}
---

```{r setup, eval = T, echo=F, warning=F,message=F, results='hide'}
library(MyFunctions)
#### Library ####
packages <- c(
  "dplyr", # Data manipulation
  "knitr",
  "kableExtra",
  "png",
  "grid",
  "janitor"
)
my_lib(packages)
```

```{r FigureS11, eval = T, echo = F, fig.cap = "Percentage change of MCP for species managed by the IPHC for end-century (2081–2100) relative to 2005–2014 under a A) high emission scenario and B) low emission scenario. Labels marked with * represent regions where models do not agree in direction of change."}
